---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# Borehole Forensics 4 - Pumping Test 2018 - 2024

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.14974957.svg)](https://doi.org/10.5281/zenodo.14974957)

<!-- badges: end -->

This dataset contains borehole forensic pumping test data, capturing key
parameters related to groundwater extraction, borehole performance, and
water quality assessments. The data was collected by BASEFlow. It includes information on the test
conditions, water levels, pumping rates, flow measurement methods, and
water quality indicators such as pH, turbidity, and electrical
conductivity.

The dataset also documents permissions granted for testing, equipment
used, and observations on borehole performance, such as whether the
water level stabilized during pumping and whether the borehole passed
the test.

Potential Uses of the Dataset 

1. Hydrogeological Analysis: Evaluating groundwater availability and borehole efficiency. 

2. Environmental Studies: Assessing water quality parameters. 

3. Infrastructure Planning: Supporting water resource management and borehole maintenance decisions.

4. Regulatory Compliance: Ensuring boreholes meet required operational standards.


## Installation

You can install the development version of boreholeforensicspumpingtest from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/boreholeforensicspumpingtest")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
library(ggplot2)
library(lubridate)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in
    your browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/boreholeforensicspumpingtest/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to the Borehole Forensics 4 - Pumping Test conducted between 2018 and 2024

```{r}
library(boreholeforensicspumpingtest)
```

### boreholeforensicspumpingtest

The dataset `boreholeforensicspumpingtest` has
`r nrow(boreholeforensicspumpingtest)` observations and `r ncol(boreholeforensicspumpingtest)`
variables

```{r}
boreholeforensicspumpingtest |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "boreholeforensicspumpingtest.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

```{r}
# Create the plot
library(boreholeforensicspumpingtest)

# Extract distinct year
year <- year(boreholeforensicspumpingtest$date_of_test)

# Create the plot
ggplot(boreholeforensicspumpingtest, aes(x = year, y = static_water_level)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Static Water Level by Year",
    x = "Year",
    y = "Static Water Level (m below reference point)"
  )


```

## License

Data are available as
[CC-BY](https://github.com/openwashdata/%7B%7B%7Bpackagename%7D%7D%7D/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("boreholeforensicspumpingtest")
```
